---
title: "RD2.qmd"
author: "Lihame Arouna" 
institute: "University of Pennyslvania" 
date: last-modified 
toc: true 

format: 
  html: 
    self-contained: true 

editor: source 
---

```{r}
library(haven)
library(tidyverse)
library(gt)
library(dplyr)
library(readxl)
library(data.table)
library(arrow) 
library(ggplot2)
library(knitr)
library(scales)
library(modelsummary)
library(lme4)
```

```{r}
AidData <- read_excel(path = "/Users/lihamearouna/Dropbox/PSCI3200/ADv3.xlsx", sheet = "GCDF_3.0") 

# Select only the necessary columns manually (update column names as needed)
AidData <- AidData[,c("Recipient", "Completion Year", "Title", "Description", "Status", "Recipient Region", 'Amount (Constant USD 2021)')]  

# Converting to data.table for memory efficiency & reloading 
AidData <- as.data.table(AidData)
write_parquet(AidData, "AidData.parquet")
AidData <- read_parquet("AidData.parquet")

#Changing Recipient name 
AidData <- AidData |> 
  rename(COUNTRY = Recipient,
         Amount = 'Amount (Constant USD 2021)', 
         Region = 'Recipient Region' ) |> 
  filter(Region == 'Africa')
head(AidData)


# Afrobarometer Merged 8 
# Defining the URL and destination file name
url <- "https://www.afrobarometer.org/wp-content/uploads/2023/03/afrobarometer_release-dataset_merge-34ctry_r8_en_2023-03-01.sav"
destfile <- "afrobarometer_r8.sav"

download.file(url, destfile = destfile, mode = "wb")
data8 <- haven::read_sav(destfile)
# Renaming numeric country column 
colnames(data8)[colnames(data8) == "COUNTRY"] <-  "country_id"

```

```{r}

#Adjusting data sets for merge 
# Create a separate data frame with country ID-country name 
country_mapping <- data.frame(
  country_id = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25,26,27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40), 
  COUNTRY = c("Angola", "Benin", "Botswana", "Burkina Faso", "Cabo Verde", "Cameroon", 
                   "CongoBrazzaville", "Côte d'Ivoire", "Eswatini", "Ethiopia", "Gabon", "Gambia", 
                   "Ghana", "Guinea", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", 
                   "Mauritania", "Mauritius", "Morocco", "Mozambique", "Namibia", "Niger", "Nigeria",
                   "São Tomé and Príncipe", "Senegal", "Seychelles", "Sierra Leone", "South Africa", 
                   "Sudan", "Tanzania", "Togo", "Tunisia", "Uganda", "Zambia", "Zimbabwe")
)
#Merging using numercic country ID to get a dataset that is traceable by country name
first_merge <- left_join(data8, country_mapping, by = "country_id")
#Grouping to match by total aid per country for sake of memory and findings 
AidData_summary <- AidData |> 
  group_by(COUNTRY) |>
  summarize(total_aid = sum(Amount, na.rm = TRUE)) 
merged_data <- left_join(AidData_summary, first_merge, by = "COUNTRY")
```

Graph 1: Plotting the relationship between total investment and public perception

```{r}
# filtering responses to only include responses of yes or no.
filtered_data <- merged_data |> filter(Q65C %in% c(0, 1))
ggplot(filtered_data, aes(x = factor(Q65C), y = total_aid)) +
  geom_boxplot(fill = "lightblue", outlier.alpha = 0.3) +
  labs(
    title = "Total Investment by Public Perception of Chinese Debt",
    x = "Public Perception (Q65C)",
    y = "Total Investment (USD)"
  ) +
  theme_bw() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_discrete(labels = c(
    `0` = "No",
    `1` = "Yes",
    `7` = "Not applicable",
    `8` = "Refused",
    `9` = "Don’t know",
    `-1` = "Missing"
  )) 

```


```{r}
# Summarize data by public perception
summary_data <- filtered_data %>%
  #calculating mean & variability of aid received by groups defined by perception of Chinese debt. 
  group_by(Q65C) %>%
  summarise(
    mean_investment = mean(total_aid, na.rm = TRUE),
    sd = sd(total_aid, na.rm = TRUE),
    n = n(),
    se = sd / sqrt(n)
  )

# Assuming your variable is Q65C
summary_data$Q65C <- factor(summary_data$Q65C, 
                         levels = c(0, 1), 
                         labels = c("No", "Yes"))

# Bar chart with error bars
ggplot(summary_data, aes(x = Q65C, y = mean_investment)) +
  geom_col(fill = "orange") +
  geom_errorbar(aes(ymin = mean_investment - se, ymax = mean_investment + se), width = 0.2) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Average Total Investment by Public Perception of Chinese Debt",
    x = "Public Perception (Q65C)",
    y = "Average Total Investment (USD)"
  ) +
  theme_minimal()
```


Covariates 
```{r}
#grouping education, employment, and trust in president data into few key groups 
analysis_data <- filtered_data |> 
  mutate(
    education_group = case_when(
      Q97 %in% 0:2 ~ "Low",
      Q97 %in% 3:5 ~ "Medium",
      Q97 %in% 6:9 ~ "High",
      TRUE ~ NA_character_
    ),
    employment_status = case_when(
      Q95A == 0 ~ "Not looking",
      Q95A == 1 ~ "Looking",
      Q95A == 2 ~ "Part-time",
      Q95A == 3 ~ "Full-time",
      TRUE ~ NA_character_
    ),
    trust_president = case_when(
      Q41A %in% c(0,1) ~ "Low",
      Q41A == 2 ~ "Medium",
      Q41A == 3 ~ "High",
      TRUE ~ NA_character_
    ),
    # Converting to factors
    education_group = factor(education_group, levels = c("Low", "Medium", "High")),
    employment_status = factor(employment_status),
    trust_president = factor(trust_president, levels = c("Low", "Medium", "High")),
    REGION = factor(REGION)
  )

# Ensuring Q65C is treated as a factor (Yes = 1, No = 0)
analysis_data$Q65C <- factor(analysis_data$Q65C, levels = c(0, 1), labels = c("No", "Yes"))

# Linear regression: Does opinion waiver based on aid received?-- Do i need this? 
model_aid <- lm(total_aid ~ Q65C + education_group + employment_status + trust_president, 
                data = analysis_data)
summary(model_aid)

# Presenting results using model summary 
modelsummary(
  model_aid,
  stars = TRUE,
  output = "markdown",
  title = "Regression: Predictors of Chinese Aid Received",
  coef_map = c(
    "Q65CYes" = "Thinks Chinese Loans Are a Problem",
    "education_groupMedium" = "Education: Medium",
    "education_groupHigh" = "Education: High",
    "employment_statusLooking" = "Employment: Looking",
    "employment_statusPart-time" = "Employment: Part-time",
    "employment_statusFull-time" = "Employment: Full-time",
    "trust_presidentMedium" = "Trust in President: Medium",
    "trust_presidentHigh" = "Trust in President: High"
  )
)

```


Empirical extension: Modeling people's opinions about Chinese debt (a binary variable, Q65C), as a function of covariates and accounting for regiohnal differences 
Estimates the log-odds that they think Chinese loans are a problem.

Adjusts estimates based on their education, employment, and trust in the president.

Adds a region-specific intercept to account for regional-level bias or influence.
```{r}

# creatihg a random effects regression model: 
random_effect_model <- glmer(
  Q65C ~ education_group + employment_status + trust_president + (1 | REGION), #estimate a seperate intercept for each region 
  data = analysis_data,
  family = binomial(link = "logit")
)

#
modelsummary(
  random_effect_model,
  stars = TRUE,
  output = "markdown",
  title = "Random Effects Logistic Regression: Predicting Views on Chinese Debt",
  coef_map = c(
    "education_groupMedium" = "Education: Medium",
    "education_groupHigh" = "Education: High",
    "employment_statusLooking" = "Employment: Looking",
    "employment_statusPart-time" = "Employment: Part-time",
    "employment_statusFull-time" = "Employment: Full-time",
    "trust_presidentMedium" = "Trust in President: Medium",
    "trust_presidentHigh" = "Trust in President: High"
  ),
  statistic = "conf.int"
)
```


Bibliography

@article{Brautigam2010, author = {Deborah Bräutigam}, title = {China's Aid Policies in Africa: Opportunities and Challenges}, journal = {The Royal Institute of International Affairs}, volume = {86}, number = {2}, pages = {401--419}, year = {2010}, doi = {10.1080/00358530903513756}, url = {https://www.tandfonline.com/doi/full/10.1080/00358530903513756} }

@report{soule2024narratives, author = {Folashadé Soulé and Cobus van Staden and Lina Benabdallah and Yu-Shan Wu and Yunnan Chen}, title = {Exploring the Role of Narratives in China-Africa Relations}, institution = {Africa Policy Research Institute}, year = {2024}, month = {June}, url = {https://afripoli.org/exploring-the-role-of-narratives-in-china-africa-relations}, doi = {10.59184/sa.044} }

@article{ChineseWesternInvestmentsAfrica2020, author = {{Peterson Najfiojor}}, title = {Chinese and Western Investments in Africa: A Comparative Analysis}, journal = {Journal of African Development}, volume = {22}, number = {2}, pages = {1--23}, year = {2020}, doi = {10.1080/21598282.2020.1846583}, url = {<https://www.tandfonline.com/doi/full/10.1080/21598282.2020.1846583%7D> }
